---
title: "Prediction of Parkinson's Disease"
author: "Group 2: Kelly Jennings, Marcus Martinez, Peter Yi, Rachel Tarrant, Stefanos Kapetanakis"
output:
  html_notebook:
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float: true
runtime: shiny
---

```{r setup, include=FALSE}
library(tidyverse)
require(MASS)
require(ggplot2)
require(ISLR)
require(data.world)
knitr::opts_chunk$set(echo = TRUE)
```
  
## **R Session Info**  

```{r}
sessionInfo()
```

## **Github Link** 
TBD

## **Connecting to data.world** 
```{r}
project <- "https://data.world/marcusgabe-ut/parkinsons-data"
data.world::set_config(cfg_env("DW_API"))
# df <- data.world::query(
#   data.world::qry_sql("SELECT * FROM parkinsons"),
#   dataset = project
# )
```

## **Introduction** 
Explanation of data... etc

## LDA
```{r}
#LDA code
```

```{r}
#LDA plots?
```

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r eruptions}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r tabsets}
shinyAppDir(
  system.file("examples/06_tabsets", package = "shiny"),
  options = list(
    width = "100%", height = 900
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.

## Inputs and Outputs 2

```{r eruptions2}
inputPanel(
  selectInput("n_breaks2", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust2", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks2),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust2)
  lines(dens, col = "blue")
})
```


